---
title: 'Isoform Project'
output:
  prettydoc::html_pretty:
    theme: cayman
    toc: true
    number_sections: true
---

# Set Up

## Load Packages
```{r message=FALSE, warning=FALSE}
library(Gviz)
library(rtracklayer)
library(Rsamtools)
library(GenomicFeatures)
library(GenomicRanges)
library(rtracklayer)
library(Rsamtools)
library(GenomicAlignments)
library(VariantAnnotation)
library(tidyverse)
library(Rsubread)
```

## Randomly Sample from Cell Subclasses
The function below will take a directory of downloaded FASTQ files (e.g. scRNAseq samples from VISp region) and create a CSV containing a random sampling of 100 file names from a certain subclass (e.g. Pvalb)

* The CSV file created should be used to specify which FASTQ files to align in STAR script (which will create BAM files)

**Parameters:**

* **filedir:** the directory where FASTQs are found (assumed to contain only files for the specified region of interest)
* **outputcsv:** the directory and desired name for random sample, e.g. "../sample.csv" (*must end with .csv*)
* **metadatacsv:** path to CSV file containing metadata for each FASTQ (should have been downloaded with FASTQ files)
* **subclass:** as a character, subclass name (defined in metadatacsv)
* **region:** as a character, region name (defined in metadatacsv)

### Examine Metadata
```{r}
## Prepare metadata
metadatacsv <- ("/external/rprshnas01/netdata_kcni/stlab/Public/AIBS_scRNAseq_2019/mouse/2020_10_24/metadata.csv")
metadata <- read_csv(metadatacsv)

## Check column names
colnames(metadata)

# See subclasses available in each region (only those with sufficient number of cells to sample)
metadata %>% filter(region_label!="HIP") %>% group_by(region_label, class_label, subclass_label) %>% tally() %>% filter(n>240)

# See list of all available regions
metadata$region_label %>% unique(.)
```

### Sample One (1) Subclass
```{r}
makerandomsample <- function(metadatacsv, filedir, subclass, region, outputcsv) {
  # Get metadata for all files
  metadata <- read_csv(metadatacsv)
  
  # Get full list of downloaded FASTQ files in region of interest
  downloaded_files <- list.files(filedir, pattern="*.fastq.tar")
  downloaded_files <- downloaded_files %>% 
    str_remove(".fastq.tar") # Removing the extension to match metadata
  
  # Of the cells in the given subclass in the given region which were downloaded, randomly sample 100
  set.seed(12345) # For reproducibility
  metadata_sample <- metadata %>% 
    filter(subclass_label == subclass) %>% 
    filter(region_label == region) %>% 
    filter(exp_component_name %in% downloaded_files) %>% 
    select(exp_component_name) %>% 
    sample_n(120)
  
  # Save the random sample to a CSV
  metadata_sample %>% 
    write_csv(outputcsv, col_names=FALSE)
}
```

### Sample Multiple Subclasses

**Additional Parameters:**

* **subclassregion:** dataframe with columns for region and subclass labels (for function: fullrandomsamples)
* **subclasses:** character vector with subclasses to be sampled (for function: makerandomsamples)
* **outputpath:** directory where .csv files will be created

```{r eval = FALSE}
# Randomly sample from each subclass in one region
## Get subclass list
subclasses <- metadata %>% 
  filter(region_label=="VISp") %>% 
  group_by(subclass_label) %>% 
  tally() %>% 
  filter(n>240) %>% 
  .$subclass_label %>% 
  str_replace_all(.," ", "_")

makerandomsamples <- function(metadatacsv, filedir, subclasses, region, outputpath){
  for (subclass in subclasses) {
    makerandomsample(metadatacsv=metadatacsv,
                     filedir=filedir,
                     subclass=subclass, region=region,
                     outputcsv=paste0(outputpath, "/", subclass, "_sample.csv"))
  }
}

# Randomly sample from each subclass in each region
## Get subclass + region dataframe
subclassregion <- metadata %>% 
  filter(region_label!="HIP" & region_label!="VISp") %>% 
  group_by(region_label, subclass_label) %>% 
  tally() %>% 
  filter(n>150) %>% 
  select(region_label, subclass_label) %>% 
  mutate(subclass_label=str_replace_all(subclass_label, " ", "_"))

fullrandomsamples <- function(metadatacsv, filedir, subclassregion, outputpath) {
  for (region in subclassregion$region_label) {
    dir.create(file.path(dirname(outputpath), region))
    outputdir <- paste0(outputpath, "/", region)
    dir.create(file.path(dirname(outputdir), "samples"))
    sampledir <- paste0(outputdir, "/samples")
    for (subclass in subclassregion$subclass_label) {
      makerandomsample(metadatacsv=metadatacsv,
                       filedir=paste0(filedir,"/",region),
                       subclass=subclass, region=region,
                       outputcsv=paste0(sampledir, "/", subclass, "_sample.csv"))
    }
  }
}
```

### Example Usage
```{r}
random100dir <- ("/external/rprshnas01/netdata_kcni/stlab/Intralab_collab_scc_projects/Isoform_project/Random100")
fastqfiles <- ("/external/rprshnas01/netdata_kcni/stlab/Public/AIBS_scRNAseq_2019/mouse/Raw")

# Create sample for L2/3 in VISp
makerandomsample(metadatacsv=metadatacsv,
                 filedir=paste0(fastqfiles, "/VISp"),
                 subclass="L2/3 IT CTX-1", region="VISp", 
                 outputcsv=paste0(random100dir, "/VISp/samples/l23_sample.csv"))


# Create samples for all subclasses in VISp
makerandomsamples(metadatacsv=metadatacsv,
                  filedir=paste0(fastqfiles, "/VISp"),
                  subclasses=subclasses, region="VISp", 
                  outputpath=paste0(random100dir, "/VISp/samples"))

# Create samples for all subclasses in all regions
fullrandomsamples(metadatacsv=metadatacsv,
                  filedir=fastqfiles,
                  subclassregion=subclassregion,
                  outputpath=random100dir)
```

## Align Sampled Cells
**TODO: MAKE VERSION THAT WORKS FOR ALL REGIONS AT ONCE**
**TODO: UPDATE SO THAT ONLY FILES WHICH ARE SUCCESSFULLY UNTARRED GO INTO SCRIPTS**

### Create Alignment Scripts
```{bash}
subclasses=('l5it' 'l5np' 'l6ct' 'l6b' 'astro')
for subclass in "${subclasses[@]}"; do
  echo "Subclass: $subclass"
  cd /external/rprshnas01/netdata_kcni/stlab/Intralab_collab_scc_projects/Isoform_project/Random100/VISp
  mkdir "$subclass"_outputs
  /external/rprshnas01/netdata_kcni/stlab/Intralab_collab_scc_projects/Isoform_project/Random100/STAR_align.sh /external/rprshnas01/netdata_kcni/stlab/Public/AIBS_scRNAseq_2019/mouse/Raw/VISp /external/rprshnas01/netdata_kcni/stlab/Intralab_collab_scc_projects/Isoform_project/Random100/VISp/${subclass}_outputs/ /external/rprshnas01/netdata_kcni/stlab/Genomic_references/Refseq/Mouse/Refseq_GRCm39/USE_THIS_genomeDir_gff /external/rprshnas01/netdata_kcni/stlab/Genomic_references/Refseq/Mouse/Refseq_GRCm39/Raw/GCF_000001635.27_GRCm39_genomic.gff /external/rprshnas01/netdata_kcni/stlab/Intralab_collab_scc_projects/Isoform_project/Random100/VISp/samples/${subclass}_sample.csv
done
```

### Run Alignment Scripts
```{bash}
subclasses=('l5it' 'l5np' 'l6ct' 'l6b' 'astro')
for subclass in "${subclasses[@]}"; do
  echo "Subclass: $subclass"
  cd /external/rprshnas01/netdata_kcni/stlab/Intralab_collab_scc_projects/Isoform_project/Random100/VISp/${subclass}_outputs
  qbatch -w 08:00:00 --ppj 12 -c 10 --mem 60G STARParaCom.txt
done
```

## Create Subclass-Level BAMs
The function below will create and index a merged BAM for a specific subclass

**Parameters:**

* **inputdir:** the directory where BAMs are found (i.e., the coord_bams directory produced from STAR script)
* **outputname:** the directory and desired name for new merged BAM, e.g. "../merged.bam" (*must end with .bam*)
* **chrID:** as a character, representing chromosome for selected reference, e.g. "NC_000082.7" for RefSeq OR "16" for Ensembl
* **coords:** as a vector, where the 1st value is the start coordinate and the 2nd is the end, e.g. c(0, 100)—note that unit is in bases, so 18,050kb becomes 18050000
* **index:** TRUE by default, change to FALSE if each BAM to be merged already has associated .bai (index file)

### For One (1) Subclass in One (1) Region
```{r}
makesubclassBAM <- function(inputdir, outputname, chrID, coords, index=TRUE) {
  
  # Create list of file paths for the BAMs
  bamfiles <- list.files(inputdir)
  bampaths <- paste0(inputdir, "/", bamfiles)
  
  # Ignoring any previously created index files
  bampaths_nobai <- str_subset(bampaths, ".bai", negate=TRUE)
  set.seed(12345) # For reproducibility
  bampaths_nobai <- bampaths_nobai %>% sample_n(100)
  
  # If index=TRUE, index each BAM file
  if (index) {
    for (file in bampaths_nobai) {
      indexBam(file)
    }
  }
  
  # Merge all BAMs into one which only contains gene of interest
  mergeBam(files=bampaths_nobai, destination=outputname, 
           overwrite=TRUE, region = GRanges(chrID, IRanges(coords[1], coords[2])), indexDestination=TRUE)
}
```

### For All Subclasses in All Regions
```{r}
makesubclassBAMs <- function(inputdir, outputdir, subclassregion, chrID, coords, index=TRUE) {
  for (region in subclassregion$region_label) {
    dir.create(file.path(dirname(outputdir), region))
    outputdir <- paste0(outputdir, "/", region)
    dir.create(file.path(dirname(outputdir), "merged_bams"))
    mergeddir <- paste0(outputdir, "/merged_bams")
    for (subclass in subclassregion$subclass_label) {
      outputname <- paste0(mergeddir, "/", "RefSeq_random100_merged_", subclass, ".bam")
      inputdir <- #TODO
      makesubclassBAM(inputdir, outputname, chrID, coords, index)
    }
  }
}
```

### Example Usage
```{r}
# For one subclass in one region
makesubclassBAM(
  inputdir=paste0(random100dir, "/VISp/l23_outputs/STAR_results/coord_bams"),
  outputname=paste0(random100dir, "/VISp/merged_bams/RefSeq_random100_merged_l23.bam"), chrID="NC_000082.7",
  coords=c(18038612,18056471), index=FALSE)

## Assign names to merged BAMs
vispsubclasses <- c("l23", "pvalb", "lamp5", "sst", "vip", "l45", "l5pt", "l6it")

for (subclass in vispsubclasses){
  assign(subclass, paste0(random100dir, "/VISp/merged_bams/RefSeq_random100_merged_", subclass, ".bam"))
}

# For all subclasses in all regions
makesubclassBAMs(inputdir=, #TODO
                 outputdir=random100dir)

## Assign names to merged BAMs
```

## Create Class-Level BAMs (TODO)
```{r}
mergeBam(files=c(l23, l45, l5pt, l6it), destination="/external/rprshnas01/netdata_kcni/stlab/Intralab_collab_scc_projects/Isoform_project/RefSeq_random100_merged_exc.bam", overwrite=TRUE, region=GRanges("NC_000082.7", IRanges(18038612,18056471)), indexDestination=TRUE)

exc <- ("/external/rprshnas01/netdata_kcni/stlab/Intralab_collab_scc_projects/Isoform_project/Random100/VISp/merged_bams/RefSeq_random100_merged_exc.bam")

mergeBam(files=c(pvalb, lamp5, sst, vip), destination="/external/rprshnas01/netdata_kcni/stlab/Intralab_collab_scc_projects/Isoform_project/RefSeq_random100_merged_inh.bam", overwrite=TRUE, region=GRanges("NC_000082.7", IRanges(18038612,18056471)), indexDestination=TRUE)

inh <- ("/external/rprshnas01/netdata_kcni/stlab/Intralab_collab_scc_projects/Isoform_project/Random100/VISp/merged_bams/RefSeq_random100_merged_inh.bam")
```

## Create Reference Genome Variables
```{r}
refseq_ref_GRCm38 <- ("/external/rprshnas01/netdata_kcni/stlab/Genomic_references/AIBS/Mouse/Refseq_GRCm38.p3/Raw/GCF_000001635.23_GRCm38.p3_genomic.gff")

refseq_ref_GRCm39 <- ("/external/rprshnas01/netdata_kcni/stlab/Genomic_references/Refseq/Mouse/Refseq_GRCm39/Raw/GCF_000001635.27_GRCm39_genomic.gff")
  
ensembl_ref_GRCm39 <- ("/external/rprshnas01/netdata_kcni/stlab/Genomic_references/Ensembl/Mouse/Release_104/Raw/Mus_musculus.GRCm39.104.gtf")
```

### Create TxDb Object
* These take a while, so you can just run the one you need

```{r results=FALSE, message=FALSE, warning=FALSE}
refseqTxDb_GRCm39 <- makeTxDbFromGFF(refseq_ref_GRCm39)

#refseqTxDb_GRCm38 <- makeTxDbFromGFF(refseq_ref_GRCm38)

#ensemblTxDb <- makeTxDbFromGFF(ensembl_ref)
```

## Set Up Figure
The function below will create a coverage + sashimi plot for two BAMs aligned to gene isoform models for the chosen reference, additionally showing coordinate positions on the x-axis

**Parameters:**

* **txdb:** a TxDb object, can be *refseqTxDb_GRCm39*, *refseqTxDb_GRCm38*, OR *ensemblTxDb*
* **bam1/bam2/...:** as a path to the BAM file (should be defined above, e.g. pvalb)
* **chrID:** as a character, representing chromosome for selected reference, e.g. "NC_000082.6" for RefSeq OR "16" for Ensembl
* **coords:** as a vector, where the 1st value is the start coordinate and the 2nd is the end, e.g. c(0, 100)—note that unit is in bases, so 18,050kb becomes 18050000
* **bam1title, etc.:** optional, a character specifying y-axis title for each coverage/junction plot

```{r}
makefigure <- function(txdb, bam1, bam2, bam3=NULL, bam4=NULL, chrID, coords, 
                       bam1title=deparse(substitute(bam1)), bam2title=deparse(substitute(bam2)), 
                       bam3title=deparse(substitute(bam3)), bam4title=deparse(substitute(bam4))) {
  
  # Extract list of transcripts
  gene_tx <- data.frame(transcripts(txdb)) %>% 
    filter(seqnames==chrID) %>% 
    filter((start>=coords[1]) & (end<=coords[2])) %>% 
    dplyr::select(tx_name) %>% 
    dplyr::filter(str_starts(tx_name, "NM")) %>% # just confirmed transcripts
    list(.$tx_name) %>% 
    .[[2]]
  
  # Create genome axis track
  genomeAxis <- GenomeAxisTrack(name="MyAxis", col="lightsteelblue4", fontcolor="lightsteelblue4", add35=TRUE) 
  
  # Create BamFile objects (used in sashimi threshold)
  bf_bam1 <- BamFile(bam1, asMates=TRUE)
  bf_bam2 <- BamFile(bam2, asMates=TRUE)
  if (!is.null(bam3) & !is.null(bam4)){
    bf_bam3 <- BamFile(bam3, asMates=TRUE)
    bf_bam4 <- BamFile(bam4, asMates=TRUE)
  }
  
  # Create plot for coverage and sashimi
  bam1_peakReads <- AlignmentsTrack(bam1, name=bam1title, cex=2, background.title="white",
                                    sashimiScore=(1/1300)*(countBam(bf_bam1)$records),
                                    col.axis="lightsteelblue4", col.title="lightsteelblue4")
  bam2_peakReads <- AlignmentsTrack(bam2, name=bam2title, cex=2, background.title="white",
                                    sashimiScore=(1/1300)*(countBam(bf_bam2)$records),
                                    col.axis="lightsteelblue4", col.title="lightsteelblue4")
  if (!is.null(bam3) & !is.null(bam4)) {
    bam3_peakReads <- AlignmentsTrack(bam3, name=bam3title, cex=2, background.title="white",
                                    sashimiScore=(1/1300)*(countBam(bf_bam3)$records),
                                    col.axis="lightsteelblue4", col.title="lightsteelblue4")
    bam4_peakReads <- AlignmentsTrack(bam4, name=bam4title, cex=2, background.title="white",
                                    sashimiScore=(1/1300)*(countBam(bf_bam4)$records),
                                    col.axis="lightsteelblue4", col.title="lightsteelblue4")
  }
  
  # Create gene models
  gr <- exonsBy(txdb, by = "tx", use.names=TRUE)[gene_tx]
  gr <- unlist(gr)
  elementMetadata(gr)$transcript <- names(gr)
  gene_models <- Gviz::GeneRegionTrack(gr, showId=TRUE, options(ucscChromosomeNames=FALSE),
                                       transcriptAnnotation="transcript", name="Gene Model",
                                       background.title="white", col.axis="lightsteelblue4",
                                       col.title="lightsteelblue4", fill="darkgrey",
                                       fontcolor.group="lightsteelblue4", col.line="lightsteelblue4")
  
  # Put the figure together
  options(ucscChromosomeNames=FALSE)
  if (!is.null(bam3) & !is.null(bam4)) {
    fig <- plotTracks(c(genomeAxis, bam1_peakReads, bam2_peakReads, bam3_peakReads, bam4_peakReads, gene_models),
             showId=TRUE,
             transcriptAnnotation="transcript",
             chromosome=chrID,
             sizes=c(1,3,3,3,3,2),
             from=coords[1],
             to=coords[2],
             extend.left=3500,
             fill="lightsteelblue", col.sashimi="lightsteelblue4",
             type=c('coverage', 'sashimi'))
  } else {
    fig <- plotTracks(c(genomeAxis, bam1_peakReads, bam2_peakReads, gene_models),
             showId=TRUE,
             transcriptAnnotation="transcript",
             chromosome=chrID,
             sizes=c(1,3,3,2),
             from=coords[1],
             to=coords[2],
             extend.left=3500,
             fill="lightsteelblue", col.sashimi="lightsteelblue4",
             type=c('coverage', 'sashimi'))
  }

  return(fig)
}
```

# Visualization: gviz

## Create Figure
* May optionally save to .tiff

### VISp: Excitatory Subclasses
```{r message=FALSE, warning=FALSE, results=FALSE}
makefigure(refseqTxDb_GRCm39, l23, l45, l5pt, l6it, "NC_000082.7", c(18038612,18056471), 
                bam1title="L2/3 IT CTX-1", bam2title="L4/5 IT CTX", bam3title="L5 PT CTX", bam4title="L6 IT CTX")
```

### VISp: Inhibitory Subclasses
```{r message=FALSE, warning=FALSE, results=FALSE}
makefigure(refseqTxDb_GRCm39, pvalb, lamp5, sst, vip, "NC_000082.7", c(18038612,18056471), 
           bam1title="Pvalb", bam2title="Lamp5", bam3title="Sst", bam4title="Vip")
```

### VISp: Excitatory vs. Inhibitory Classes
```{r message=FALSE, warning=FALSE, results=FALSE}
#tiff(filename="pvalbvsl23.tiff", units="in", width=8, height=6, res=500)
makefigure(refseqTxDb_GRCm39, bam1=exc, bam2=inh, chrID="NC_000082.7", coords=c(18038612,18056471), 
           bam1title="Excitatory", bam2title="Inhibitory")
#dev.off()
```

# Quantification (still testing)

## See exon boundary coordinates for every transcript
* Parameters same as defined above

```{r}
getexoncoords <- function(txdb, chrID, coords) {
  gene_tx <- data.frame(transcripts(txdb)) %>% 
    filter(seqnames==chrID) %>% 
    filter((start>=coords[1]) & (end<=coords[2])) %>% 
    dplyr::select(tx_name) %>% 
    list(.$tx_name) %>% 
    .[[2]]
  gr <- exonsBy(txdb, by = "tx", use.names=TRUE)[gene_tx]
  df <- data.frame(unlist(gr)) %>% 
    mutate(transcript=str_match(exon_name, "-(.*)-")[,2]) %>% 
    mutate(exon_num=str_match(exon_name, ".*-(.*)")[,2]) %>% 
    select(transcript, exon_num, start, end)
  return(df)
}
```

### Example Usage
```{r}
exoncoords <- getexoncoords(refseqTxDb_GRCm39, "NC_000082.7", c(18038612,18056471))
exoncoords
```

## Define transcript-specific ranges
Select exon coordinates unique to one transcript (exon in one tx2 should be a subset of exon in tx1)
```{r}
selectrange <- function(chrID, exoncoords, tx1, tx2, exon1, exon2) {
  exon1_rng <- exoncoords %>% filter(transcript==tx1 & exon_num==exon1)
  exon2_rng <- exoncoords %>% filter(transcript==tx2 & exon_num==exon2)
  gr_tx1 <- GRanges(chrID, IRanges(exon1_rng$start, exon1_rng$end))
  gr_tx2 <- GRanges(chrID, IRanges(exon2_rng$start, exon2_rng$end))
  diff <- GenomicRanges::setdiff(gr_tx1, gr_tx2)
  
  return(diff)
}
```

### Example Usage
```{r}
rng <- selectrange("NC_000082.7", exoncoords, "NM_172151.4", "NM_001379019.1", "11", "11")
```

## summarizeOverlaps: calculate normalized reads

### As a percentage of reads for whole gene
```{r}
#Pvalb cells
bf_pvalb <- BamFile(pvalb, asMates=TRUE)

counts_pvalb <- data.frame(union=assays(summarizeOverlaps(rng, bf_pvalb, singleEnd=FALSE, fragments=TRUE))$counts)
colSums(counts_pvalb) # raw number of reads in region
p_pvalb <- 100*(rowSums(counts_pvalb))/(countBam(bf_pvalb)$records) # reads as proportion of total reads in gene

#L2/3
bf_l23 <- BamFile(l23, asMates=TRUE)

counts_l23 <- data.frame(union=assays(summarizeOverlaps(rng, bf_l23, singleEnd=FALSE, fragments=TRUE))$counts)
colSums(counts_l23)
p_l23 <- 100*(rowSums(counts_l23))/(countBam(bf_l23)$records)

colSums(data.frame(p_pvalb))
colSums(data.frame(p_l23))
```

### Region Ratios (short vs long-minus-short)
```{r}
long <- rng
short <- GRanges("NC_000082.7", IRanges(18038612, 18039568))

long_pvalb <- data.frame(union=assays(summarizeOverlaps(long, bf_pvalb, singleEnd=FALSE, fragments=TRUE))$counts)
short_pvalb <- data.frame(union=assays(summarizeOverlaps(short, bf_pvalb, singleEnd=FALSE, fragments=TRUE))$counts)

long_l23 <- data.frame(union=assays(summarizeOverlaps(long, bf_l23, singleEnd=FALSE, fragments=TRUE))$counts)
short_l23 <- data.frame(union=assays(summarizeOverlaps(short, bf_l23, singleEnd=FALSE, fragments=TRUE))$counts)

ratio_pvalb <- (rowSums(long_pvalb))/(rowSums(short_pvalb))
ratio_l23 <- (rowSums(long_l23))/(rowSums(short_l23))

readratios <- data.frame(Pvalb=ratio_pvalb, "L2/3 IT"=ratio_l23, row.names="Read Ratio")
```

### As a percentage of reads in long final exon region
```{r}
full <- GRanges("NC_000082.7", IRanges(18038617, 18041241))
full_minus_short <- GRanges("NC_000082.7", IRanges(18039568, 18041241))

full_pvalb <- data.frame(union=assays(summarizeOverlaps(full, bf_pvalb, singleEnd=FALSE, fragments=TRUE))$counts)
full_minus_short_pvalb <- data.frame(union=assays(summarizeOverlaps(full_minus_short, bf_pvalb, singleEnd=FALSE, fragments=TRUE))$counts)

full_l23 <- data.frame(union=assays(summarizeOverlaps(full, bf_l23, singleEnd=FALSE, fragments=TRUE))$counts)
full_minus_short_l23 <- data.frame(union=assays(summarizeOverlaps(full_minus_short, bf_l23, singleEnd=FALSE, fragments=TRUE))$counts)

p_pvalb <- 100*(rowSums(full_minus_short_pvalb))/(rowSums(full_pvalb))
p_l23 <- 100*(rowSums(full_minus_short_l23))/(rowSums(full_l23))

readpercentages <- data.frame(Pvalb=p_pvalb, "L2/3 IT"=p_l23, row.names="Read %")
```

## summarizeJunctions
18041241-18042082 (long final exon) 
* Pvalb = 50
* L23: 16

18039568-18042082 (short final exon) 
* Pvalb = 8
* L23: 117

```{r}
galign_pvalb <- readGAlignmentPairs(pvalb,index="/external/rprshnas01/netdata_kcni/stlab/Intralab_collab_scc_projects/Isoform_project/RefSeq_random100_merged_pvalb.bam", strandMode=1)
galign_l23 <- readGAlignmentPairs(l23, index="/external/rprshnas01/netdata_kcni/stlab/Intralab_collab_scc_projects/Isoform_project/RefSeq_random100_merged_l23.bam", strandMode=1)

junctions_pvalb <- summarizeJunctions(galign_pvalb)
junction_df_pvalb <- data.frame(c(as.data.frame(junctions_pvalb@ranges), as.data.frame(junctions_pvalb$score)))

junctions_l23 <- summarizeJunctions(galign_l23)
junction_df_l23 <- data.frame(c(as.data.frame(junctions_l23@ranges), as.data.frame(junctions_l23$score)))

junction_df_l23 <- junction_df_l23 %>% filter(end==18042081 & start!=18033262)
junction_df_pvalb <-junction_df_pvalb %>% filter(end==18042081)

junc_ratio_pvalb <- junction_df_pvalb$junctions_pvalb.score[2]/junction_df_pvalb$junctions_pvalb.score[1]
junc_ratio_l23 <- junction_df_l23$junctions_l23.score[2]/junction_df_l23$junctions_l23.score[1]

junc_percent_pvalb <- 100*junction_df_pvalb$junctions_pvalb.score[2]/(junction_df_pvalb$junctions_pvalb.score[1]+junction_df_pvalb$junctions_pvalb.score[2])
junc_percent_l23 <- 100*junction_df_l23$junctions_l23.score[2]/(junction_df_l23$junctions_l23.score[1]+junction_df_l23$junctions_l23.score[2])

junctionratios<- data.frame(Pvalb=junc_ratio_pvalb, "L2/3 IT"=junc_ratio_l23, row.names="Junction Ratio")
junctionpercentages<- data.frame(Pvalb=junc_percent_pvalb, "L2/3 IT"=junc_percent_l23, row.names="Junction %")
```

## Comparing Metrics
```{r}
rbind(readratios,junctionratios,readpercentages,junctionpercentages)
```
